---
title: "Pratical Machine Learning Project"
author: "Fernando Henrique Macedo"
date: "20 de junho de 2016"
output: html_document
---

# Load libraries and data

## Libraries
```{r libraries, message=FALSE}
library(caret)
library(randomForest)
```

## Data
```{r data}
trainSet <- read.csv("pml-training.csv")
testSet <- read.csv("pml-testing.csv")
```

# Cleaning data

The trainSet contains `r ncol(trainSet)` columns. First, we should clean the columns with lots of NA or blank values.

```{r clean NA}
NA_cols <- sapply(trainSet, function(x) sum(is.na(x) | x == ""))
table(NA_cols)
```

The line above gave us an array with the number of non-NA values per column. Looking at the unique values of it, there are two possible values (`r unique(NA_cols)`), what shows a lot of columns with only NA values. So, we should remove them.

```{r remove NA}
trainClean <- trainSet[, NA_cols == 0]
dim(trainClean)
```

# Feature selection

Observing the first rows of the cleaned data:

```{r firts rows}
head(trainClean)
```


The first 7 columns seem irrelevant as features.

```{r feature}
trainFinal <- trainClean[, -(1:7)]
```

# Spliting data

```{r splitting}
set.seed(25361)
splt_idx <- createDataPartition(trainFinal$classe, p = 0.7, list = FALSE)
training <- trainFinal[splt_idx,]
testing <- trainFinal[-splt_idx,]
```

# Modeling

We will use a Random Forest with 10-fold validation to improve accuracy.

```{r randomforest}
trainingModel <- train(classe ~ ., data = training, method = "rf",
                       trControl = trainControl(method = "cv", number = 10),
                       ntree = 251)
trainingModel
```

Now, we could observe the accuracy of the model in a confusion matrix

```{r confusion}
testPrediction <- predict(trainingModel, testing)
confusionMatrix(testing$classe, testPrediction)
```

So, we have an accuracy of `r confusionMatrix(testing$classe, testPrediction)$overall[[1]]`.

# Predicting the Test Data

```{r pred test}
testData <- testSet[, which(NA_cols == 0)]
testData <- testData[, -c(1:7, ncol(testData))]
predict(trainingModel, testData)
```

